Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)

### 2024-01-18 manuscript/revision2/revision2.tex
manuscript: revisions to text from SAS from 20240110, some small additional revisions from AHM

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally-cued phenological responses: for example, populations of
sunflower (Helianthus annuus) exhibit day-neutral, facultative short
day, and facultative long-day flowering responses, which vary with their
environments (5, 6). Distinct genetic responses in different
environments are known as genotype by environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in a number
of cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE. Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or sign-changing GxE, at the level of
individual loci (22--25), previous work has found limited evidence of
antagonistic pleiotropy (12, 26). However, this work has been limited by
a known statistical bias that reduced detection of genetic effects that
differ in sign (26--28). Thus, despite substantial interest in the
frequencies of various forms of GxE, the prevalence of antagonistic
pleiotropy relative to other forms of GxE remains unknown.

@@

Here, we estimate the ways genetic effects on phenological timings
depend on weather prior to the phenological event. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define a number of ways phenological timings might covary with
weather (@tbl-covar) and additional ways phenological timings might vary
by site (@SI-info), then jointly re-estimate genetic effects on these
timings at all eight sites using the set of these covariance types that
significantly improved the modeled log-likelihood when included
(@SI-info) (35). The method we use for effect re-estimation, mash,
allows us to identify and specify multiple ways genetic effects may
covary with the environment, and does not have a statistical bias in
detecting genetic effects with the same or opposite signs (36). To
confirm our genetic mapping of GxWeather, we compare our re-estimated
genetic effects to mapping results from an outbred pseudo-F2 cross grown
at the same sites. Our analysis allow us to describe the weather cues
and genetic variation affecting phenology in two divergent natural
populations of switchgrass.

@@

Genotypes from the Gulf and Midwest subpopulations had distinct
phenological timings and distinct patterns of phenological correlations
across our eight common garden sites (Figure~\ref{fig-map}). At the
three Texas common gardens (hereafter `Texas' gardens), located within
the natural range of the Gulf subpopulation, Gulf vegetative growth
occurred before Midwestern vegetative growth, and Gulf flowering
occurred after Midwestern flowering (Figure~\ref{fig-map} A). At the
four northernmost common gardens (hereafter `North' gardens), located
within the natural range of the Midwest subpopulation, both Gulf
vegetative growth and flowering occurred after Midwest vegetative growth
and flowering. At the Oklahoma common garden, located near the natural
range limits of both the Gulf and the Midwest subpopulations, Gulf and
Midwest vegetative growth occurred over the same time period, and Gulf
flowering occurred after Midwestern flowering (Figure~\ref{fig-map} A).
These patterns led to strong negative phenotypic correlations for the
timing of vegetative growth between the North and Texas gardens,
particularly in the Gulf and across all individuals (hereafter, `Both'
subpopulations), and contributed to positive phenotypic correlations for
the timing of flowering that had larger magnitudes at more northern
gardens (Figure~\ref{fig-map} B).

@@

\subsubsection{GxWeather mixtures for vegetative and reproductive
timing}\label{gxweather-mixtures-for-vegetative-and-reproductive-timing}

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using a model that
allows effects to covary between gardens in many ways
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). To do this, we used
genome-wide association (GWAS) on site-specific BLUPs for the timing of
vegetative growth and flowering to determine preliminary single
nucleotide polymorphism (SNP) effect estimates at each garden. We then
jointly modeled genetic effect estimates, and the mixture of ways these
effects might covary, across all eight common gardens, using a two-part
procedure. We first modeled genetic effects with different combinations
of covariance types, using a subset of effects selected at random from
all SNPs used in the GWAS. In these models, we specified many ways
genetic effects could covary between gardens (as expanded on below) and
used a greedy algorithm to iteratively select covariance structures that
significantly improved the log likelihood of a model when included.
Given the high correlation between some of the covariance patterns we
tested, this approach allowed us to select only the subset of patterns
of covariation that most improved the fit of the joint effect patterns
present in the data. We then re-estimated the set of genetic effects
that had the largest effect estimates in univariate GWAS (hereafter
``strong'' effects) using the set of significant covariance matrices
that most improved the model fit on the random effects.

We created three categories of ways effects could covary across gardens:
``canonical'' covariance, with simple patterns of effects;
``data-driven'' covariances derived from common patterns of SNP effects
in the data, and ``GxWeather'' covariances that captured the covariance
of weather variables at specific times before the phenological event
(Table 1, SI Appendix, Section S2). For example, GxWeather covariance
between gardens in MI and MO, for rainfall in the seven days prior to
flowering, captures the covariance in the amount of rainfall genotype A,
B, and C receive in the seven days prior to their three flowering dates
in MI and their three flowering dates in MO. If these six rainfall
values are similar, covariance will be near one; if they are
uncorrelated, covariance will be near zero; if there is sufficient rank
changing in rainfall amounts between gardens for these genotypes, this
covariance will be near negative one.

The GxWeather covariances allow hypothesis testing of specific weather
variables as cues for each genetic effect. Say that a SNP in FLC
controls flowering in a photoperiod-dependent manner. In that case, that
SNP effect could have a high mixture proportion, or mass, on a
covariance matrix created using a photoperiod-based environmental cue,
such as day length two days prior to flowering. In our data, we would
infer that the effect of that SNP on flowering was caused by a response
to that or a correlated environmental cue. Overall, loadings of genetic
effects on these GxWeather matrices in the model with re-estimated
effects provide information on genome-wide patterns of SNP-environment
interaction.

The phenotypic correlations for the timing of vegetative growth had
moderate negative correlations between the Texas and North gardens,
particularly in the Gulf and in Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, then they might be controlled by SNP effects that
differed in sign across these regions (effects with antagonistic
pleiotropy), SNP effects that are large in one garden or region and
non-significant in others (effects with amplification), or a combination
of these patterns. Five GxWeather covariance matrices were selected by
the greedy mash algorithm, one to two per subpopulation. Of these five,
three had mass on them in mash models of the strong effects
(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
covariances between sites between Texas and North gardens
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances. SNP-associated phenotypic effects covaried with different
weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix of daylength 14 days prior to the date of vegetative
growth. The covariance matrix for this weather cue was very similar to
the pattern of phenotypic correlation for the timing of vegetative
growth in the Gulf subpopulation (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). Mash models of the timing of vegetative
growth in Midwest and Both subpopulations did not include this GxWeather
covariance; the Midwest had no weight on any GxWeather covariance, while
Both subpopulations had non-zero weights on two additional GxWeather
covariance types, average temperature one day prior to green-up, and the
day length change in seconds in the day prior to green-up
(Figure~\ref{fig-covar} B). The average temperature covariance matrix
had negative covariances between Texas and North gardens, though not as
strong as the negative phenotypic correlations seen in Both
subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
the Gulf subpopulation and Both subpopulations had mass on any GxWeather
covariance matrices (Figure~\ref{fig-covar} C).

For flowering date, distinct GxWeather covariances captured
SNP-associated effect patterns in the Gulf and Midwest subpopulations.
33\% of SNP effects on flowering in the Gulf subpopulation covaried with
cumulative rainfall in the seven days prior to flowering (Fig 2D).
22.6\% of SNP effects on flowering in the Midwest subpopulation covaried
with day length change in the two days prior to flowering (Fig 2D),
which had negative covariances between Texas and North gardens. Neither
covariance matrix was selected for in Both subpopulations, and no
GxWeather covariance matrices had mass in Both subpopulations (Fig 2E).
In five of the six mash models of strong effects, the GxWeather
covariance matrices captured a minority of the posterior weights of the
strong effects (Figure 2C,E); the majority of this mass was on various
canonical covariance matrices. These matrices included simple
heterozygosity, with intermediate, positive covariances between all
gardens, and garden-specific effects present only at one garden.

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. We used the local
false sign rate (lfsr), an analogue of the lfdr that establishes
confidence in the effect sign, not the effect's difference from zero, to
determine significance. We required significance (p \textless{} 0.05) in
both gardens to include effects. This means that our tests for
antagonistic pleiotropy, or a sign change between conditions, carry an
equal statistical burden to those for effects with the same sign. We
also summarized the pairwise patterns of effect sign and magnitude
within and between the Texas and North regions (Table 2).

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens (Fig. 3A). 78.7\%
of pairwise comparisons between North and Texas gardens had a difference
in sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
comparisons had a difference in sign, respectively (Table 2). The
majority of pairwise effects for greenup for the Midwest
(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
same sign, and effects frequently differed in magnitude between the MO
and OK garden and other gardens (Table 2; Fig 3A).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Table 2). More effects differed in
magnitude between the Texas and North regions than within these regions
(42.7\% vs \textless20\%; Figure 3B). The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign (Table 2). Most
differences in sign were between TX1, the southernmost garden, and all
other gardens (Fig 3B). Similarly, more effects that differed in
magnitude included gardens in the Texas region (52.3-55.9\%), and most
effect pairs in the North region were not distinguishable (91.5\%).

@@

We sought additional experimental support for our significant effect
re-estimates using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites (Fig.
4A,B). We conducted quantitative trait loci (QTL) mapping of flowering
as functions of four environmental cues that we also used as covariance
matrices in mash, and identified eight QTL for flowering date, six QTL
for flowering GDD, ten QTL for flowering day length, and eight QTL
flowering day length change, all of which showed QTL by environment
interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7). All flowering QTL
intervals contained at least one SNP significant in at least one mash
run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
tail of significance, whichever was stricter (SI Appendix, Dataset 8).
We also looked for enrichments of mash SNPs in the 1\% tail of
significance (the `mash 1\% tail') within each QTL interval. At the 5\%
level, five QTL had enrichments of SNPs in the mash 1\% tail. Overall,
there were eleven significant enrichments (p \textless{} 0.05,
hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
Our QTL intervals had more enrichments of SNPs in the mash 1\% tail than
were found for all but eleven of these sets of random genomic intervals
(Fig. 4C, p = 0.011). Thus, we were able to experimentally support our
significant effect re-estimates with a QTL mapping experiment using a
separate mapping population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (Figure 3A). This result supports
theoretical models that local adaptation should involve antagonistic
pleiotropy at the level of individual loci (22--25), and is the first
experimental work using QTL mapping and GWAS across common gardens to
find antagonistic pleiotropy to be common in small genomic regions (12,
37, 38).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal in the two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on photoperiod. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. The genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (Fig. 4C).

>>>>

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (Helianthus annuus) exhibit day-neutral, facultative short
day, and facultative long-day flowering responses, which vary with their
environments (5, 6). Distinct genetic responses in different
environments are known as genotype by environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE. Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or sign-changing GxE, at the level of
individual loci (22--25), previous work has found limited evidence of
antagonistic pleiotropy (12, 26). However, this work has been limited by
a known statistical bias that reduced detection of genetic effects that
differ in sign (26--28). Thus, despite substantial interest in the
frequencies of various forms of GxE, the prevalence of antagonistic
pleiotropy relative to other forms of GxE remains unknown.

@@

Here, we estimate the ways genetic effects on phenological timings
depend on weather prior to the phenological event. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(@tbl-covar) and additional ways phenological timings might vary by site
(@SI-info), then jointly re-estimate genetic effects on these timings at
all eight sites using the set of these covariance types that
significantly improved the modeled log-likelihood when included
(@SI-info) (35). We use the Bayesian framework \emph{mash} (multivariate
adaptive shrinkage), developed by (35), to refine effect size estimates
from genome-wide association (GWAS) conducted on individuals at each of
the eight sites. \emph{mash} allows us to identify and specify multiple
covariance structures among genetic effect estimates across sites,
including structures that represent covariance in weather variables of
interest. Importantly, \emph{mash} does not have a statistical bias in
detecting genetic effects with the same or opposite signs (36). To
confirm our genetic mapping of GxWeather, we compare our posterior
genetic effect estimates from \emph{mash} to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
to describe the weather cues and genetic variation affecting phenology
in two divergent natural populations of switchgrass.

@@

Genotypes from the Gulf and Midwest subpopulations had distinct
phenological timings and distinct patterns of phenological correlations
across our eight common garden sites (Figure~\ref{fig-map}). At the
three Texas common gardens (hereafter `Texas' gardens), located within
the natural range of the Gulf subpopulation, Gulf vegetative growth
occurred before Midwestern vegetative growth, and Gulf flowering
occurred after Midwestern flowering (Figure~\ref{fig-map} A). At the
four northernmost common gardens (hereafter `North' gardens), located
within the natural range of the Midwest subpopulation, both Gulf
vegetative growth and flowering occurred after Midwest vegetative growth
and flowering. At the Oklahoma common garden, located near the natural
range limits of both the Gulf and the Midwest subpopulations, Gulf and
Midwest vegetative growth occurred over the same period, and Gulf
flowering occurred after Midwestern flowering (Figure~\ref{fig-map} A).
These patterns led to strong negative phenotypic correlations for the
timing of vegetative growth between the North and Texas gardens,
particularly in the Gulf and across all individuals (hereafter, `Both'
subpopulations), and contributed to positive phenotypic correlations for
the timing of flowering that had larger magnitudes at more northern
gardens (Figure~\ref{fig-map} B).

@@

\subsubsection{\texorpdfstring{Inference of GxWeather effects for
vegetative and reproductive timing using
\emph{mash}}{Inference of GxWeather effects for vegetative and reproductive timing using mash}}\label{inference-of-gxweather-effects-for-vegetative-and-reproductive-timing-using-mash}

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which allows effects to covary between gardens in many ways
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). To do this, we used
GWAS on site-specific BLUPs for the timing of vegetative growth and
flowering to determine preliminary single nucleotide polymorphism (SNP)
effect estimates for both traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using a subset of randomly selected SNPs used in the GWAS
and their corresponding effect size estimates. In these models
(hereafter referred to as ``random''), we then specified a variety of
potential covariance structures among genetic effect size estimates from
each garden (as expanded on below) and used a greedy algorithm to
iteratively select and add covariance structures that significantly
improved the log likelihood of the \emph{mash} model. Given the high
correlation between some of the covariance patterns we tested, this
approach allowed us to select only the subset of patterns of covariance
that significantly improved the fit of the jointly inferred structure of
genetic effect estimates present in the data. We then selected the set
of SNPs and corresponding effect size estimates that had the largest
effect estimates in univariate GWAS (hereafter ``strong'' effects) and
jointly re-estimated these effect size estimates at all eight sites
using the set of significant covariance matrices that most improved the
model fit on the random effects.

We specified three categories of covariance structures across gardens:
``canonical'' covariance, with simple patterns of effect size covariance
introduced in the initial \emph{mash} manuscript; ``data-driven''
covariances derived from common patterns of SNP effects observed in the
data, and ``GxWeather'' covariance, estimated from the covariance of
empirical weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example,
GxWeather covariance between gardens in MI and MO, for rainfall in the
seven days prior to flowering, captures the covariance in the amount of
rainfall genotypes A, B, and C receive in the seven days prior to their
three flowering dates in MI and their three flowering dates in MO. If
these six rainfall values are similar, covariance will be near one; if
they are uncorrelated, covariance will be near zero; if there is
sufficient rank changing in rainfall amounts between gardens for these
genotypes, this covariance will be near negative one. We specified
multiple covariance structures for the canonical, data-driven, and
GxWeather categories, each of which represented a specific pattern of
covariance in genetic effect size estimates that could be included in
the model.

The GxWeather covariance structures allow hypothesis testing of specific
weather variables as cues for the start of vegetative and reproductive
growth. Say that a SNP in FLC controls flowering in a
photoperiod-dependent manner. In that case, the joint estimate of
effects for that SNP could have a high mixture proportion, or mass, on a
covariance matrix created using a photoperiod-based environmental cue,
such as day length two days prior to flowering. In our data, we would
infer that the effect of that SNP on flowering was caused by a response
to the environmental cue used to construct the GxWeather covariance
structure with the largest mass. Overall, loadings of genetic effects on
these GxWeather matrices in the model, along with the posterior effect
size estimates provide information on genome-wide patterns of
SNP-environment interaction.

The phenotypic correlations for the timing of vegetative growth had
moderate negative correlations between the Texas and North gardens,
particularly in the Gulf subpopulation and Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, then they may be controlled by SNP effects that
differed in sign across these regions (effects with antagonistic
pleiotropy), SNP effects that are large in one garden or region and
non-significant in others (effects with amplification), or a combination
of these patterns. Five GxWeather covariance structures were selected by
the greedy algorithm, one to two per subpopulation. Of these five, three
had mass on them in mash models of the strong effects
(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
covariances between sites between Texas and North gardens
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances. SNP-associated phenotypic effects covaried with different
weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix constructed using the covariance of daylength 14 days
prior to the date of vegetative growth. The covariance matrix for this
weather cue was very similar to the pattern of phenotypic correlation
for the timing of vegetative growth in the Gulf subpopulation
(Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash models of the
timing of vegetative growth in the Midwest subpopulation and Both
subpopulations did not include this GxWeather covariance; the Midwest
had no weight on any GxWeather covariance, while Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). Only the Gulf subpopulation and Both
subpopulations had mass on any GxWeather covariance matrices
(Figure~\ref{fig-covar} C).

For flowering date, distinct GxWeather covariance structures captured
covariance in effect sizes for SNPs in the Gulf and Midwest
subpopulations. 33\% of SNP effects on flowering in the Gulf
subpopulation covaried with cumulative rainfall in the seven days prior
to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
flowering in the Midwest subpopulation covaried with day length change
in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
negative covariances between Texas and North gardens. Neither covariance
matrix was selected as significantly improving the log-likelihood of the
\emph{mash} model of Both subpopulations, and no GxWeather covariance
matrices had a non-zero mass in models of Both subpopulations
(Figure~\ref{fig-covar} E). In five of the six mash models of strong
effects, the GxWeather covariance matrices captured a minority of the
posterior weights of the strong effects (Figure~\ref{fig-covar} C,E);
the majority of this mass was on various canonical covariance matrices.
These matrices included simple heterozygosity, with intermediate,
positive covariances between all gardens, and garden-specific effects
present only at one garden.

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. We used the local
false sign rate (lfsr), an analogue of the lfdr that establishes
confidence in the effect sign, not the effect's difference from zero, to
determine significance. We required lfsr significance (p \textless{}
0.05) in both gardens to include effects. This means that our tests for
antagonistic pleiotropy, or a sign change between conditions, carry an
equal statistical burden to those for effects with the same sign. We
also summarized the pairwise patterns of effect sign and magnitude
within and between the Texas and North regions, within the natural
ranges of the Gulf and Midwest subpopulations (@fig-effects).

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens (@fig-effects A).
78.7\% of pairwise comparisons between North and Texas gardens had a
difference in sign, while only 28.6\% and 0.2\% of North-North or
Texas-Texas comparisons had a difference in sign, respectively
(@fig-effects). The majority of pairwise effects for greenup for the
Midwest (\textgreater55\%) and Both (\textgreater85\%) subpopulations
were the same sign, and effects frequently differed in magnitude between
the MO and OK garden and other gardens (@fig-effects X).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (@fig-effects). More effects differed in
magnitude between the Texas and North regions than within these regions
(42.7\% vs \textless20\%; @fig-effects X). The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign (@fig-effects). Most
differences in sign were between TX1, the southernmost garden, and all
other gardens (@fig-effects B). Similarly, more effects that differed in
magnitude included gardens in the Texas region (52.3-55.9\%), and most
effect pairs in the North region were not distinguishable (91.5\%).

@@

We sought additional experimental support for our joint re-estimates of
SNP effects using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites (@fig-qtl
A,B). We conducted quantitative trait loci (QTL) mapping of flowering as
functions of four environmental cues that we also used as covariance
matrices in mash, and identified eight QTL for flowering date, six QTL
for flowering GDD, ten QTL for flowering day length, and eight QTL
flowering day length change, all of which showed QTL by environment
interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7). All flowering QTL
intervals contained at least one SNP significant in at least one mash
run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
tail of significance, whichever was stricter (SI Appendix, Dataset 8).
We also looked for enrichments of mash SNPs in the 1\% tail of
significance (the `mash 1\% tail') within each QTL interval. At the 5\%
level, five QTL had enrichments of SNPs in the mash 1\% tail. Overall,
there were eleven significant enrichments (p \textless{} 0.05,
hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
Our QTL intervals had more enrichments of SNPs in the mash 1\% tail than
were found for all but eleven of these sets of random genomic intervals
(@fig-qtl C, p = 0.011). Thus, we were able to experimentally support
our joint re-estimates of SNP effects with a QTL mapping experiment
using a separate mapping population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (@fig-effects A). This result
supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (22--25), and is
the first experimental work using QTL mapping and GWAS across common
gardens to find antagonistic pleiotropy to be common in small genomic
regions (12, 37, 38).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal in the two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on photoperiod. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. The genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (@fig-qtl C).
